Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2012-10-16 paper.tex
fixed silly abs error
minor paper edits.

\begin{algorithm}[t] \label{alg:loglinear}
\SetAlCapFnt{\small}
\SetAlTitleFnt{}
\caption{Mackenthun's algorithm with pilot symbols}
\DontPrintSemicolon
\KwIn{$\{y_i, i \in P \cup D \}$}
\For{$i \in D$ \nllabel{alg_loop_setup}}{
$\phi = \angle{y_i}$ \;
$u = \round{\phi} $ \;
$z_i = \phi -  u $ \;
$g_i = y_i e^{-j u}$ \;
}
$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_sortindices} \;
$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
$\hat{a} = \frac{1}{L} Y$ \;
$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
$\eta = e^{j2\pi/M} - 1$ \;
\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
$Y = Y + \eta g_{\sigma(k)}$ \;
$g_{\sigma(k)} = (\eta + 1) g_{\sigma(k)} $\;
$Q = \frac{1}{L}\abs{Y}^2$\;
\If{$Q > \hat{Q}$}{
 	$\hat{Q} = Q$ \;
 	$\hat{a} =  \frac{1}{L} Y$ \;
 }
}
\Return{$\hat{a}$ \nllabel{alg_return}}
\end{algorithm}

@@

We will have particular use of complex random variables of the form $1 + W$ where $W$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
\[
R e^{j\Phi} = 1 + W.
\]
The joint pdf of $R$ and $\Phi$ can be shown to be
\begin{equation}\label{eq:pdfRPhi}
f(r,\phi) = \frac{r f_Z(\sqrt{\abs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\abs{r^2 - 2r\cos\phi + 1}}}.
\end{equation}
Since $\cos\phi$ has period $2\pi$ and is even on $[-\pi,\pi]$ it follows that $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$.  The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.  So,
\begin{equation}\label{eq:expectRepartRphi}
\expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
\end{equation}
where $\Re(\cdot)$ denotes the real part, and
\begin{equation}\label{eq:expectImpartRphi}
\expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
\end{equation}
where $\Im(\cdot)$ denotes the imaginary part.

@@

\begin{theorem}\label{thm:consistency} (Almost sure convergence)
Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}^2$ finite, and let $\{y_i, i \in P \cup D\}$ be given by~\eqref{eq:sigmod}.   Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$. %where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.  
Put $L = \abs{P \cup D}$ and let $p$ and $d$ be the limits
\[
\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d \qquad \text{as $L \rightarrow \infty$.}
\] 
Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
\begin{equation}\label{eq:RiandPhii}
R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i} ,
\end{equation}
and define the continuous function
\[
G(x) = p h_1(x) + d h_2(x) \qquad \text{where,}
\]
\[
h_1(x) = \expect R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \expect R_1 \cos\sfracpart{ x + \Phi_1}.
\]
If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then:
\begin{enumerate}
\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
\end{enumerate}
\end{theorem}

@@

A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.  This assumption asserts that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $\abs{G(x_i)} \rightarrow \sabs{G(0)}$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.  Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.  The next proposition describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.  The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.  The proof is given in Appendix~\ref{appendix:prop2proof}.

\begin{proposition}\label{prop:contgg2}
Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having continuous and differentiable joint pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.  Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.  Let,
\[
h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{x + \Phi}.
\]
Then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
\end{proposition}

@@

Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect to the data symbols,
 \begin{align*}
SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
\end{align*}
where
\[
Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
\]
and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.  Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$, 
\begin{equation}\label{eq:hatrhoZ}
\hat{\rho} = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
\end{equation}
where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
\[
SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
\]

We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.  Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
\begin{align*}
y_i &= a_0 s_i + w_i \\
&= a_0 s_i \left( 1 + \frac{w_i}{a_0 s_i} \right) \\
&= a_0 s_i R_i e^{j \Phi_i} \\
&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
\end{align*}
%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta},
\begin{align*}
\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
&\equiv \round{ \fracpart{\theta_0 - \theta}_{\pi} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
&= \round{ \lambda + \Phi_i + \angle{s_i} },
\end{align*}
where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
\begin{align}
 y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \nonumber \\
&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \nonumber  \\
&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}} \label{eq:yethetadhat}
\end{align}
since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
\[
y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
\]
Now,
\[
Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
\]
and
\[
\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
\] 
Let 
\begin{equation}\label{eq:GLdefn}
G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
\end{equation}
and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.  Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.  We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.  The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.

Recall the functions $G$, $h_1$ and $h_2$ defined in the statement of Theorem~\ref{thm:consistency}.  Observe that
\[
\expect G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
\]
and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
\[
\lim_{L \rightarrow\infty} \expect G_L(\lambda) = G(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
\]
As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.  Let $A$ be the subset of the sample space $\Omega$ upon which $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ as $L\rightarrow\infty$.  Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.  Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.  Because $\sabs{G(x)}$ is uniquely maximised at $x=0$, it follows that $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$. So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.  Part 1 of Theorem~\ref{thm:consistency} follows.

@@

We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.  Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.  Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.  The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.  This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.  However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$. Define the function
\begin{equation}\label{eq:RLdef}
R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
\end{equation}
Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.  Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
\begin{align*}
0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
\end{align*}
Lemma~\ref{lem:Qconv} shows that
\[
\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
\]
where $o_P(1)$ denotes a number converging in probability to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.  Lemma~\ref{lem:empiricprocc} shows that
\[
\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
\]
It follows from the three equations above that,
\[
0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
\]
and rearranging gives,
\[
\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
\]
Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Theorem~\ref{thm:normality}.  It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
\[
\frac{pA_1 + dA_2}{(p + Hd)^2}.
\]

@@

We present the results of Monte-Carlo simulations with the least squares estimator.  In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.  Under these conditions the least squares estimator is also the maximum likelihood estimator.  Simulations are run with $M=2,4,8$ (BPSK and QPSK and $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.  The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.  For each value of SNR, $T = 5000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.

>>>>

\begin{algorithm}[t] \label{alg:loglinear}
\SetAlCapFnt{\small}
\SetAlTitleFnt{}
\caption{Mackenthun's algorithm with pilot symbols}
\DontPrintSemicolon
\KwIn{$\{y_i, i \in P \cup D \}$}
\For{$i \in D$ \nllabel{alg_loop_setup}}{
$\phi = \angle{y_i}$ \;
$u = \round{\phi} $ \;
$z_i = \phi -  u $ \;
$g_i = y_i e^{-j u}$ \nllabel{alg_gset} \;
}
$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_sortindices} \;
$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
$\hat{a} = \frac{1}{L} Y$ \nllabel{alg_hata1} \;
$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
$\eta = e^{j2\pi/M} - 1$ \;
\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
$Y = Y + \eta g_{\sigma(k)}$ \;
$g_{\sigma(k)} = (\eta + 1) g_{\sigma(k)} $\;
$Q = \frac{1}{L}\abs{Y}^2$\;
\If{$Q > \hat{Q}$}{
 	$\hat{Q} = Q$ \;
 	$\hat{a} =  \frac{1}{L} Y$ \nllabel{alg_hata2} \;
 }
}
\Return{$\hat{a}$ \nllabel{alg_return}}
\end{algorithm}

@@

We will have particular use of complex random variables of the form $1 + W$ where $W$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
\[
R e^{j\Phi} = 1 + W.
\]
The joint pdf of $R$ and $\Phi$ can be shown to be
\begin{equation}\label{eq:pdfRPhi}
f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
\end{equation}
Since $\cos\phi$ has period $2\pi$ and is even on $[-\pi,\pi]$ it follows that $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$.  The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.  So,
\begin{equation}\label{eq:expectRepartRphi}
\expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
\end{equation}
where $\Re(\cdot)$ denotes the real part, and
\begin{equation}\label{eq:expectImpartRphi}
\expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
\end{equation}
where $\Im(\cdot)$ denotes the imaginary part.

@@

\begin{theorem}\label{thm:consistency} (Almost sure convergence)
Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}^2$ finite, and let $\{y_i, i \in P \cup D\}$ be given by~\eqref{eq:sigmod}.   Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$. %where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.  
Put $L = \abs{P \cup D}$ and let $p$ and $d$ be the limits
\[
\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d \qquad \text{as $L \rightarrow \infty$.}
\] 
Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
\begin{equation}\label{eq:RiandPhii}
R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i} ,
\end{equation}
and define the continuous function
\[
G(x) = p h_1(x) + d h_2(x) \qquad \text{where,}
\]
\[
h_1(x) = \expect R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \expect R_1 \cos\sfracpart{ x + \Phi_1}.
\]
If $p > 0$ and if $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then:
\begin{enumerate}
\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
\end{enumerate}
\end{theorem}

@@

A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.  This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.  Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.  The next proposition describes a class of circularly symmetric distributions for which $G(x)$ is uniquely maximised at $x = 0$.  The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.  The proof is given in Appendix~\ref{appendix:prop2proof}.

\begin{proposition}\label{prop:contgg2}
Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having continuous and differentiable joint pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.  Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.  Let,
\[
h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{x + \Phi}.
\]
Then $h_1(x)$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $h_2(x)$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
\end{proposition}

@@

Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect to the data symbols,
 \begin{align*}
SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
\end{align*}
where
\[
Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
\]
and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.  Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$ conditioned on minimisation with respect to the data symbols, 
\begin{equation}\label{eq:hatrhoZ}
\hat{\rho}(\theta) = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
\end{equation}
where $\Re(\cdot)$ denotes the real part.  Since $\hat{\rho}(\theta)$ is nonnegative it follows that $\Re(Z(\theta))$ is nonnegative for all $\theta \in [-\pi, \pi)$.  Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
\[
SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
\]

We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\Re(Z(\theta))$.  Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
\begin{align*}
y_i &= a_0 s_i + w_i \\
&= a_0 s_i \left( 1 + \frac{w_i}{a_0 s_i} \right) \\
&= a_0 s_i R_i e^{j \Phi_i} \\
&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
\end{align*}
%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta},
\begin{align*}
\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
&\equiv \round{ \fracpart{\theta_0 - \theta}_{\pi} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
&= \round{ \lambda + \Phi_i + \angle{s_i} },
\end{align*}
where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
\begin{align}
 y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \nonumber \\
&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \nonumber  \\
&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}} \label{eq:yethetadhat}
\end{align}
since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
\[
y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
\]
Now,
\[
Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
\]
and
\[
\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
\] 
Let 
\begin{equation}\label{eq:GLdefn}
G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
\end{equation}
and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.  Since $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ then  $\hat{\lambda}_L$ is the maximiser of $G_L(\lambda)$.  We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.  The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.

Recall the functions $G$, $h_1$ and $h_2$ defined in the statement of Theorem~\ref{thm:consistency}.  Observe that
\[
\expect G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
\]
and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
\[
\lim_{L \rightarrow\infty} \expect G_L(\lambda) = G(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
\]
As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.  Let $A$ be the subset of the sample space $\Omega$ upon which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.  Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.  Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.  Because $G(x)$ is uniquely maximised at $x=0$, it follows that $G(\hat{\lambda}_L) \rightarrow G(0)$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$. So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.  Part 1 of Theorem~\ref{thm:consistency} follows.

@@

We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.  Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.  Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.  The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.  This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.  However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$. Define the function
\begin{equation}\label{eq:RLdef}
R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
\end{equation}
Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$, and so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.  Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
\begin{align*}
0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
\end{align*}
Lemma~\ref{lem:Qconv} shows that
\[
\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
\]
where $o_P(1)$ denotes a number converging in probability to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.  Lemma~\ref{lem:empiricprocc} shows that
\[
\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
\]
It follows from the three equations above that,
\[
0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
\]
and rearranging gives,
\[
\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
\]
Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Theorem~\ref{thm:normality}.  It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
\[
\frac{pA_1 + dA_2}{(p + Hd)^2}.
\]

@@

We present the results of Monte-Carlo simulations with the least squares estimator.  In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.  Under these conditions the least squares estimator is also the maximum likelihood estimator.  Simulations are run with $M=2,4,8$ (BPSK and QPSK and $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB} in steps of \unit[1]{dB}.  The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.  For each value of SNR, $T = 5000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
